### 2019-04-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Started to write the « related works for other models » section in Chapter:5
J'avance. En train de sauver notre borne inférieure.

% -----------------------------------------------------------------
\section{Motivation for multi-player MAB models}
\label{sec:5:introduction}
% -----------------------------------------------------------------

@@

For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$).
With the presence of a central controller that can assign the devices to separate channels, this amounts to choosing at each time step \emph{several} arms of a MAB in order to maximize the global rewards, and can thus be viewed as an application of the multiple-play bandit, introduced by \cite{Anantharam87a} and recently studied by \cite{Komiyama15}.
They essentially proved that existing algorithms can be easily extended to the multiple-play case, with provable guarantees on their regret.
% In the context of CR, this first extension can model a centralized agent taking decisions for end-devices, for instance a base station affecting devices to channels.
% Multiple-Play MAB algorithms successfully used for other Cognitive Radio tasks,
% for instance by \cite{Maghsudi16} for 5G small-cells or by \cite{Bonnefoi16} for Internet-of-Things (IoT) networks.
%
% Another generalization targets the opposite goal, where several devices
% are learning and conjointly accessing the same network,
% trying to find a consensus in a decentralized learning way.
% It is harder as there is a game theoretic-like situation, for instance where devices try to avoid collisions.
%
Due to the communication cost implied by a central controller, a more relevant model is the
\emph{decentralized multi-player} multi-armed bandit model, introduced by \cite{Zhao10} and \cite{Anandkumar10,Anandkumar11}, in which players select arms individually and collisions may occur, that yield a loss of reward.
Further algorithms were proposed in similar models by \cite{Tekin12IEEE} and \cite{Kalathil12} (under the assumption that each arm is a Markov chain)
and by \cite{Avner15,Avner16} and \cite{Rosenski16} (for \iid{} or piece-wise \iid{} arms).
%
The goal for every player is to select most of the time one of the $M$ best arms, without colliding too often with other players.
A first difficulty relies in the well-known trade-off between \emph{exploration} and \emph{exploitation}: players need to explore all arms to estimate their means while trying to focus on the best arms to gain as much rewards as possible.
The decentralized setting considers no exchange of information between players, that only know $K$ and $M$, and to avoid collisions, players should furthermore find orthogonal configurations (\ie, the $M$ players use the $M$ best arms without any collision), without communicating.
Hence, in that case the trade-off is to be found between exploration, exploitation \emph{and} low collisions.

All these above-mentioned works are motivated by the OSA problem, in which it is assumed that \emph{sensing} occurs, that is each smart device observes the availability of a channel (sample from the arm) \emph{before} trying to transmit and possibly experiment a collision with other smart devices.
However some real radio networks do not use sensing at all, \eg, emerging standards developed for \emph{Internet of Things} (IoT) networks such as LoRaWAN.
Thus, to take into account these new applications, algorithms with additional constraints on the available feedback have to be proposed within the multiple-player MAB model.
Especially, the typical approach that combines a (single-player) bandit algorithm based on the sensing information --to learn the quality of the channels while targeting the best ones-- with a low-complexity decentralized collision avoidance protocol, is no longer possible.

In this chapter, we take a step back and present the different feedback levels possible for multi-player MAB algorithms. For each of them, we propose algorithmic solutions supported by both experimental and theoretical guarantees. In the presence of sensing information, our contributions are a new problem-dependent regret lower bound, tighter than previous work, and the introduction of two algorithms, \RandTopM{} and \MCTopM{}. Both are shown to achieve an asymptotically optimal number of selections of the sub-optimal arms, and for \MCTopM{} we furthermore establish a logarithmic upper bound on the regret, that follows from a careful control of the number of collisions. In the absence of sensing information, we propose the \Selfish{} heuristic and investigate its performance. Our study of this algorithm is supported by (promising) empirical performance and some first (disappointing) theoretical elements.

@@

% -----------------------------------------------------------------
\section{Multi-player bandit model with different feedback levels}
\label{sec:5:model}
% -----------------------------------------------------------------

% describe our stochastic assumptions
We consider a $K$-armed Bernoulli bandit model, % (for $K\geq2$),
in which arm $k$ is a Bernoulli distribution with mean $\mu_k\in[0,1]$.
We denote $(Y_{k,t})_{t\in\N}$ the \iid{} (binary) \emph{reward stream} for arm $k$, that satisfies $\Pr(Y_{k,t}=1) = \mu_k$ and that is independent from the other rewards streams.
However we mention that our lower bound and all our algorithms (and their analysis) can be easily extended to one-dimensional exponential families (just like for the \klUCB{} algorithm of \cite{KLUCBJournal}). For simplicity, we focus on the Bernoulli case, that is also the most relevant for Cognitive Radio, as it can model channel availabilities.

@@

Note that other models for rewards loss have also been proposed in the literature, for instance the reward is randomly allocated to one of the players selecting it.
To stay consistent with the model presented in the previous Chapter~\ref{chapter:4},
we preferred to focus on full reward occlusion in this chapter.

A multi-player MAB strategy is a tuple $\rho = (\rho^1,\dots,\rho^M)$ of arm selection strategies for of each of the $M$ players, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
First, each player $j$ should adopt a \emph{sequential} strategy $\rho^j$, that decides which arm to select at time $t$ based on \emph{previous observations}.
Previous observations for player $j$ at time $t$ always include the previously chosen arms $A^j(s)$ and received rewards $r^j(s)$ for $s<t$, but may also include the \emph{sensing information} $Y_{A^j(t),t}$ or the \emph{collision information} $C^j(t)$.
More precisely, depending on the application, one may consider the following three observation models, \modelun, \modeldeux{} and \modeltrois.

@@

\begin{itemize}
  \item[\modelun]
    \textbf{Simultaneous sensing and collision}: player $j$ observes  $Y_{A^j(t),t}$ \emph{and} $C^j(t)$ (not previously studied).
  \item[\modeldeux]
    \textbf{Sensing, then collision}: player $j$ observes $Y_{A^j(t),t}$, \emph{then} observes the reward, and thus also $C^j(t)$ only if $Y_{A^j(t),t} = 1$.
    This common setup, studied for example by \cite{Anandkumar11,Avner15,Rosenski16}, is relevant to model the OSA problem: the device first checks for the presence of primary users in the chosen channel.
    If this channel is free ($Y_{A^j(t),t}=1$), the transmission is successful ($r^j(t)=1$) if no collision occurs with other smart devices ($\overline{C^j(t)}$).
  \item[\modeltrois]
    \textbf{No sensing}: player $j$ only observes the reward $r^j(t)$.
    For IoT networks, this reward can be interpreted as an acknowledgement from a Base Station,
    received when a communication was successful.
    A lack of acknowledgment may be due to a collision
    with a device from the background traffic $(Y_{A^j(t),t}=0)$,
    or to a collision with one of the others players ($C^j(t)$).
    However, the sensing and collision information are censored.
    Recently, our work presented in Chapter~\ref{chapter:4} \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications. %\footnote{\cite{Bonnefoi17} considers a different setting, more suited for large-scale IoT networks, and a future work will be to extend our work to this setting with more than $K$ players but small probability of activations.}.
\end{itemize}

@@

% FIXME
\begin{framed}
  \paragraph{\textcolor{red}{Warning}.}
  The regret lower bound given in \cite{Besson2018ALT} is not applicable to any algorithm,
  as it was discovered by Etienne Boursier and Vianney Perchet, as they explain in Section~2.4 in \cite{BoursierPerchet18}.
  The proof we gave in the Appendix of our paper \cite{Besson2018ALT} was wrong in just one step,
  but the results stated in the section are now correct,
  thanks to the modifications proposed by Emilie Kaufmann and Abbas Mehrabian in Appendix~E of \cite{KaufmannAbbas19}.
  The cost for this modification is to restrict the lower-bound to a smaller class of algorithms, which contains both \RhoRand{} and our two proposals \RandTopM{} and \MCTopM{} but not the two algorithms proposed SIC-MMAB from \cite{BoursierPerchet18} and Multiplayer Explore-then-Commit (M-ETC) from \cite{KaufmannAbbas19}.
\end{framed}
% FIXME

@@

To express our lower bound, we need to introduce $\kl(x,y)$ as the Kullback-Leibler divergence between the Bernoulli distribution of mean $x$ and that of mean $y$, so that $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$.
%
We first introduce the assumption under which we derive a regret lower bound, that generalizes a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.

@@

We now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections under a \emph{decentralized strategy} that has access to the sensing information.
This result, proved in Appendix~\ref{proof:5:BetterLowerBound}, yields an asymptotic logarithmic lower bound on the regret, also given in Theorem~\ref{thm:5:BetterLowerBound}.

\begin{theorem}\label{thm:5:BetterLowerBound}
  Under observation models $\modelun$ and $\modeldeux$, for any strongly uniformly efficient \emph{decentralized} policy $\rho$ and $\boldsymbol{\mu}\in\cP_M$,
  % \vspace*{-10pt}  % XXX
  \begin{equation}\label{eq:5:LBDraws}
    \forall j \in \{1,\dots,M\}, \ \forall k \in \Mworst, \ \ \ \liminf_{T\to \infty} \frac{\E_{\mu}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k, \mu_M^*)}.
  \end{equation}

>>>>

% -----------------------------------------------------------------
\section{Motivations for multi-player MAB models}
\label{sec:5:introduction}
% -----------------------------------------------------------------

@@

For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$), as we explained in Chapters~\ref{chapter:1,chapter:4}.
With the presence of a central controller that can assign the devices to separate channels, this amounts to choosing at each time step \emph{several} arms of a MAB in order to maximize the global rewards, and can thus be viewed as an application of the multiple-play bandit, introduced by \cite{Anantharam87a} and recently studied by \cite{Komiyama15}.
They essentially proved that existing algorithms can be easily extended to the multiple-play case, with provable guarantees on their regret.
% In the context of CR, this first extension can model a centralized agent taking decisions for end-devices, for instance a base station affecting devices to channels.
% Multiple-Play MAB algorithms successfully used for other Cognitive Radio tasks,
% for instance by \cite{Maghsudi16} for 5G small-cells or by \cite{Bonnefoi16} for Internet-of-Things (IoT) networks.
%
% Another generalization targets the opposite goal, where several devices
% are learning and conjointly accessing the same network,
% trying to find a consensus in a decentralized learning way.
% It is harder as there is a game theoretic-like situation, for instance where devices try to avoid collisions.
%
Due to the communication cost implied by a central controller, a more relevant model is the
\emph{decentralized multi-player} multi-armed bandit model, introduced by \cite{Zhao10} and further studied shortly after in \cite{Anandkumar10,Anandkumar11}, in which players select arms individually and collisions may occur, that yield a loss of reward.
Further algorithms were proposed in similar models by \cite{Tekin12IEEE} and \cite{Kalathil12} (under the assumption that each arm is a Markov chain)
and by \cite{Avner15,Avner16} and \cite{Rosenski16} (for \iid{} or piece-wise \iid{} arms).
%
The goal for every player is to select most of the time one of the $M$ best arms, without colliding too often with other players.
A first difficulty relies in the well-known trade-off between \emph{exploration} and \emph{exploitation}: players need to explore all the arms to estimate their means, while trying to focus on the best arms to gain as much rewards as possible.
The decentralized setting considers wireless protocol with no direct or explicit exchange of information between players, and assumes that the players only know $K$ and $M$. To avoid collisions, players should furthermore find orthogonal configurations, \ie, the $M$ players use the $M$ best arms without any collision, without explicitly communicating\footnote{One must now be careful about this aspect when stating that ``no explicit communications are allowed between players'' in the model, as \cite{BoursierPerchet18} proved that even in the no sensing case, the ``communication trick'' can be used to exchange information between players, by generating collisions at some pre-agreed times. We discuss this work more in details at the end of this chapter, in Section~\ref{sub:5:withoutSensing}.} with each other.
Hence, in that case the trade-off is to be found between exploration, exploitation \emph{and} low collisions.

All these above-mentioned works are motivated by the OSA problem, in which it is assumed that \emph{sensing} occurs, that is each smart device observes the availability of a channel (sample from the arm) \emph{before} trying to transmit and possibly experiment a collision with other smart devices.
However some real radio networks do not use sensing at all, \eg, emerging standards currently or recently developed for \emph{Internet of Things} (IoT) networks, such as LoRaWAN.
Thus, to take into account these new applications, algorithms with additional constraints on the available feedback have to be proposed within the multiple-player MAB model.
Especially, the typical approach that combines a (single-player) bandit algorithm based on the sensing information --to learn the quality of the channels while targeting the best ones-- with a low-complexity decentralized collision avoidance protocol, is no longer possible.
Our article \cite{Besson2018ALT} was the first to study this other model of multi-player bandits for the ``no sensing'' case.

In this chapter, we take a step back and present the different feedback levels possible for multi-player MAB algorithms. For each of them, we propose algorithmic solutions supported by both experimental and theoretical guarantees. In the presence of sensing information, our contributions are a new problem-dependent regret lower bound, tighter than previous work, and the introduction of two algorithms, \RandTopM{} and \MCTopM{}. Both are shown to achieve an asymptotically optimal number of selections of the sub-optimal arms, and for \MCTopM{} we furthermore establish a logarithmic upper bound on the regret, that follows from a careful control of the number of collisions. In the absence of sensing information, we propose the \Selfish{} heuristic and investigate its performance. Our study of this algorithm was supported in \cite{Besson2018ALT} by (promising) empirical performance and some first (disappointing) theoretical elements, and it was proved to be inefficient since then.

@@

% -----------------------------------------------------------------
\section{Three feedback levels for the multi-player bandit model}
\label{sec:5:model}
% -----------------------------------------------------------------

% describe our stochastic assumptions
Similarly to what is presented in Chapter~\ref{chapter:2},
we consider a $K$-armed Bernoulli bandit model, % (for $K\geq2$),
in which arm $k$ is a Bernoulli distribution with mean $\mu_k\in[0,1]$.
We denote $(Y_{k,t})_{t\in\N}$ the \iid{} (binary) \emph{reward stream} for arm $k$, that satisfies $\Pr(Y_{k,t}=1) = \mu_k$ and that is independent from the other rewards streams.

\textbf{Only Bernoulli?}
However we mention that our lower bound and all our algorithms (and their analysis) can be easily extended to one-dimensional exponential families (just like for the \klUCB{} algorithm of \cite{KLUCBJournal}). For simplicity, we focus on the Bernoulli case, that is also the most relevant for Cognitive Radio, as it can model channel availabilities.

@@

Note that other models for rewards loss have also been proposed in the literature, for instance the reward is randomly allocated to one of the players selecting it.
To stay consistent and for simplicity, with the model presented in the previous Chapter~\ref{chapter:4},
we preferred to focus on full reward occlusion in this chapter.

A multi-player MAB strategy is formally defined as a tuple $\rho = (\rho^1,\dots,\rho^M)$ of arm selection strategies for of each of the $M$ players, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
First, each player $j$ should adopt a \emph{sequential} strategy $\rho^j$, that decides which arm to select at time $t$ based on \emph{previous observations}.
Previous observations for player $j$ at time $t$ always include the previously chosen arms $A^j(s)$ and received rewards $r^j(s)$ for $s<t$, but may also include the \emph{sensing information} $Y_{A^j(t),t}$ or the \emph{collision information} $C^j(t)$.
More precisely, depending on the application, one may consider the following three observation models, \modelun, \modeldeux{} and \modeltrois.

@@

\begin{itemize}
  \item[\modelun]
    \textbf{Simultaneous sensing and collision}: player $j$ observes  $Y_{A^j(t),t}$ \emph{and} $C^j(t)$.
    We note that this first model was never previously studied, but we do not focus on it because of its unrealistic aspect, and its simplicity (it is easier than the following model, for which we obtain good theoretical results).
  \item[\modeldeux]
    \textbf{Sensing, then collision}: player $j$ observes $Y_{A^j(t),t}$, \emph{then} observes the reward, and thus also $C^j(t)$ only if $Y_{A^j(t),t} = 1$.
    This common setup, studied for example by \cite{Anandkumar11,Avner15,Rosenski16}, is relevant to model the OSA problem: the device first checks for the presence of primary users in the chosen channel.
    If this channel is free ($Y_{A^j(t),t}=1$), the transmission is successful ($r^j(t)=1$) if no collision occurs with other smart devices ($\overline{C^j(t)}$).
  \item[\modeltrois]
    \textbf{No sensing}: player $j$ only observes the reward $r^j(t)$.
    For IoT networks, this reward can be interpreted as an acknowledgement from a Base Station,
    received when a communication was successful.
    A lack of acknowledgment may be due to a collision
    with a device from the background traffic $(Y_{A^j(t),t}=0)$,
    or to a collision with one of the others players ($C^j(t)$).
    However, the sensing and collision information are censored.
    Recently, our work presented in Chapter~\ref{chapter:4} \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications, but as explained above we did not present theoretical results for this third model (yet).
    %\footnote{\cite{Bonnefoi17} considers a different setting, more suited for large-scale IoT networks, and a future work will be to extend our work to this setting with more than $K$ players but small probability of activations.}.
\end{itemize}

@@

% FIXME
\begin{framed}
  \textbf{\textcolor{red}{Warning}.}
  The regret lower bound given in \cite{Besson2018ALT} is not applicable to any algorithm,
  as it was discovered by Etienne Boursier and Vianney Perchet, as they explain in Section~2.4 in \cite{BoursierPerchet18}.
  The proof we gave in the Appendix of our paper \cite{Besson2018ALT} was wrong in just one step,
  but the results stated in the section are now correct,
  thanks to the modifications proposed by Emilie Kaufmann and Abbas Mehrabian in Appendix~E of \cite{KaufmannAbbas19}.
  The cost for this modification is to restrict the lower-bound to a smaller class of algorithms, which contains both \RhoRand{} and our two proposals \RandTopM{} and \MCTopM{} but not the two algorithms proposed SIC-MMAB from \cite{BoursierPerchet18} and Multiplayer Explore-then-Commit (M-ETC) from \cite{KaufmannAbbas19}.
\end{framed}
% FIXME

@@

To express our lower bound, we need to introduce $\kl(x,y)$ as the Kullback-Leibler divergence between the Bernoulli distribution of mean $x$ and that of mean $y$, so that $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$.
%
We first introduce the assumption under which we derive a regret lower bound, that generalizes a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.
Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
This class contains our proposals, \RandTopM{} and \MCTopM, but rule out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.

@@

As it was discovered in \cite{BoursierPerchet18} and further explained in Appendix~E of \cite{KaufmannAbbas19}, our lower-bound was not true for any algorithm, as we wrongly claimed in \cite{Besson2018ALT} that the collision information term is zero for any algorithm $\rho$.
However, we wanted to be still able to present our derivations and results giving the desired lower-bound, mainly because its proof and its form give insights on how to design an asymptotically optimal algorithm $\rho$, and because it still applies to algorithms of the \RhoRand{} family as well as our proposals \RandTopM{} and \MCTopM.
The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} come from Appendix~E of \cite{KaufmannAbbas19}.
% FIXME
% We let the interested reader fetch the details on their paper.
Fix a player $j\in\{1,\dots,M\}$, and we consider the observations $\cO_t$ gathered by this player after $t$ rounds of its algorithm $\rho^j$:
\[ \cO_t := \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right),\]
where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$ (\eg, for an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index, the decision is an $\argmax$ that is usually a uniform random selection among the arms with maximum index).
Fix a problem $\bm{\mu}$, and as we do in Appendix~\ref{proof:5:BetterLowerBound}, introduce an alternative model parametrized by $\bm{\lambda}$, with a small difference between $\bm{\mu}$.

For the algorithm $\rho$, we introduce the \emph{collision information term} as:
\begin{equation}\label{eq:5:collisionInformationTerm}
  \cI_{\bm{\mu},\bm{\lambda}}(\rho,T) := \sum_{t=1}^{T} \mathrm{KL}(\P_{\bm{\mu}}^{C_t | \cO_{t-1}}, \P_{\bm{\lambda}}^{C_t | \cO_{t-1}}).
\end{equation}

We now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections under a \emph{decentralized strategy} $\rho$ that has access to the sensing information and satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T) = \smallO{\log(T)}$.
This additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T)$ essentially says that ``the collisions do not bring too much information on the arm means'', as stated in \cite{KaufmannAbbas19}.
This result, proved in Appendix~\ref{proof:5:BetterLowerBound}, yields an asymptotic logarithmic lower bound on the regret, also given in Theorem~\ref{thm:5:BetterLowerBound}.

\begin{theorem}\label{thm:5:BetterLowerBound}
  Under observation models $\modelun$ and $\modeldeux$, for any strongly uniformly efficient \emph{decentralized} policy $\rho$ and $\boldsymbol{\mu}\in\cP_M$.
  Furthermore, if $\rho$ satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T) = \smallO{\log(T)}$, then
  % \vspace*{-10pt}  % XXX
  \begin{equation}\label{eq:5:LBDraws}
    \forall j \in \{1,\dots,M\}, \ \forall k \in \Mworst, \ \ \ \liminf_{T\to \infty} \frac{\E_{\mu}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k, \mu_M^*)}.
  \end{equation}

### 2019-04-08 2-Chapters/5-Chapter/chapter5.tex
J'avance. En train de sauver notre borne inférieure.

In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
We consider $M$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
A perfect time and frequency synchronization is assumed,
and some \iid{} background traffic is assumed to be non-uniformly in the $K$ channels.
As before, if two or more devices decide to use the same channel at the same time, a \emph{collision} arises and none of sent uplink packet can be received by the gateway.
%
Such networks can be modelled using a decentralized multi-player multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
Unfortunately, it is very hard to formally analyze the IoT network model we presented in Chapter~\ref{chapter:4}, mainly because of the random activation process of all the dynamic learning devices, Even if there is many identical bandit algorithms learning independently and in a decentralized way, the difficulty mainly comes from the fact that all of them are only communicating at some (random) time steps, and at each time step even the number of communicating devices is random and varies a lot.

Mainly for these two reasons, for this Chapter~\ref{chapter:5} we prefer to only consider the easier case of at most $M \leq K$ devices, communicating at each time step.
Each device will use a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward at each time step.
%
We start by reviewing previous works on multi-player MAB models, which all considered the easier case of \emph{sensing feedback}.
Each time frame is separated as before in a \emph{sensing} phase (during which the device senses for the background traffic),
an \emph{uplink} phase (during which the device sends a packet to the gateway if it sensed the chosen channel to be free),
and a \emph{downlink} phase (during which it waits for an \Ack{} from the gateway).
In this first model, the binary reward is $1$ only if the channel was sensed to be free of background traffic \textbf{and} if \Ack{} was received, and the device has access to both information.
For this first model, we present two algorithms, \RandTopM{} and \MCTopM, based on the combination of an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
Like in previous works, we consider the centralized system regret (multi-player regret), or simply referred to as regret.
We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions..
We then analyze our \MCTopM{} algorithm and we show that its regret upper-bound asymptotically matches the improved lower-bound, with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
We also present extensive numerical simulations that show that our proposal outperforms all previous solutions and is much more efficient in this easier model of sensing feedback with a fixed and known number of players $M$ accessing $K \geq M$ channels.

We are also interested to consider the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we find on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can has a linear regret with a low probability (and so asymptotically it has linear regret and fails to solve the no sensing multi-player MAB problem).
We did not propose any other efficient algorithm, but our work presented in April~$2018$ inspired some other articles \cite{LugosiMehrabian18,BoursierPerchet18}.
Two articles confirmed our finding that \Selfish{} can have a linear regret, and they all proposed new algorithms.
We not include some numerical simulations to compare some of them, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.

>>>>

In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
We consider $M$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
A perfect time and frequency synchronization is assumed,
and some \iid{} background traffic is assumed to be non-uniformly in the $K$ channels.
As before, if two or more devices decide to use the same channel at the same time, a \emph{collision} arises and none of sent uplink packet can be received by the gateway.
%
Such networks can be modelled using a decentralized multi-player multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
Unfortunately, it is very hard to formally analyze the IoT network model we presented in Chapter~\ref{chapter:4}, mainly because of the random activation process of all the dynamic learning devices.
Even if there are many identical bandit algorithms learning independently and in a decentralized way, the difficulty mainly comes from the fact that all of them are only communicating at some (random) time steps, and at each time step the number of communicating devices is random and unpredictable.

Mainly for these two reasons, for this Chapter~\ref{chapter:5} we prefer to only consider the easier case of at most $M \leq K$ devices, communicating at each time step.
Each device will use a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each uplink message (\ie, at each time step).
%
We start by reviewing previous works on multi-player MAB models, which all considered the easier case of \emph{sensing feedback}.
Each time frame is separated as before in a \emph{sensing} phase (during which the device senses for the background traffic),
an \emph{uplink} phase (during which the device sends a packet to the gateway if it sensed the chosen channel to be free),
and a \emph{downlink} phase (during which it waits for an \Ack{} from the gateway).
In this first model, the binary reward is $1$ only if the channel was sensed to be free of background traffic \textbf{and} if \Ack{} was received, and the device has access to both information.
For this first model, we present two algorithms, \RandTopM{} and \MCTopM, based on the combination of an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
Like in previous works, we consider the centralized system regret (multi-player regret), or simply referred to as regret.
We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
We then analyze our \MCTopM{} algorithm and we show that its regret upper-bound is logarithmic and order-optimal, improving over the previous state-of-the-art.
%  with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
We also present extensive numerical simulations that show that our proposal outperforms all previous solutions and is much more efficient in this easier model of sensing feedback with a fixed and known number of players $M$ accessing $K \geq M$ channels.

We are also interested in the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we found on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can has a linear regret with a low probability (and so asymptotically it has linear expected regret and fails to solve ``no sensing'' multi-player MAB problems).
We did not propose any other efficient algorithm, but our work presented in April~$2018$ inspired some other articles \cite{LugosiMehrabian18,BoursierPerchet18}.
These two articles confirmed our finding that \Selfish{} can have a linear regret, and they both proposed new algorithms.
We include some numerical simulations to compare some of them, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
